---
title: "Classification_Muyodi_Based on polygon"
author: "Ronald Reagan Okoth"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages needed for this session
```{r}
library(raster)
library(sf)
library(sp)
library(rgdal)
library(dplyr)
library(tibble)
library(randomForest)
library(rasterVis)
library(ggplot2)

```

- Define path to the rastyer images and the training data
- Read the csv file containing training data,
- Transform it to a spatial object and reproject it to the projection of the raster files

```{r}
raster_path <-  '/Users/rragankonywa/OneDrive/Moyodi_classification/Data/Muyodi_images_2019/'
train_data_path <- '/Users/rragankonywa/OneDrive/Moyodi_classification/Data/training_data_clean4.shp'

# read the csv file
train_data <- st_read(train_data_path)[1]
# grab the crs of one of the raster files
raster_crs <- crs(stack(paste0(raster_path, list.files(raster_path)[1])))
#crs(train_data)
#crs(train_data)

```

 Order the data based on the Name column then
 Create a column in gthe training data with unique class id for each class
```{r}
train_data <- train_data[order(train_data$Name), ]
id = list()
all_Names <- train_data$Name
unique_Names <- unique(all_Names)

for(i in 1:length(unique_Names)){
  x <- all_Names[all_Names==unique_Names[i]] 
  id <- append(id,rep(i, length(x))) %>% 
    unlist()
}


train_data <- train_data %>% 
  add_column(id, .before="Name")

train_data
```


Overlay the training polygpns on an RGB image to visualize the data a bit
Normaly, one would need to make sure the crs of the shapefile and that of the raster are similayr for the overlay to work properly
```{r}
sample_raster <- stack(paste0(raster_path, list.files(raster_path)[1]))
plotRGB(sample_raster, r=4,g=3,b=2, scale= 65535, stretch= "lin")
plot(train_data[1], add = TRUE, lwd = 2, ext = extent(sample_raster), border = "red")
train_data <- st_zm(train_data)

```

For Each Year (only working with 2019 fir this case) read each image, subset only bands (blue, green, red and NIR)
and stack the bands in a single image with many bands (nbands = 4*the number of images u have foe a particulare year)
```{r}
band_subset <- c(2,3,4,8) 
raster_files <- list.files(raster_path, pattern = ".tif")
final_stack <- stack()
bandnames <- c("B", "G", "R", "NIR", "ndvi")
#bandnames <- c(paste0("B",1:12), "ndvi")
for(i in 1:length(raster_files)) {
  bandname <- unlist(lapply(bandnames, function(each){
    paste0(each,"_", "img", i)
  }))
  raster_image <- stack(paste0(raster_path, raster_files[i]))[[c(2,3,4,8)]]
  ndvi <- (raster_image[[4]]-raster_image[[3]])/(raster_image[[4]]+raster_image[[3]])
  #ndvi <- (raster_image[[8]]-raster_image[[4]])/(raster_image[[8]]+raster_image[[4]])
  raster_with_ndvi <- stack(raster_image,ndvi)
  names(raster_with_ndvi) <- bandname
  
  final_stack <- stack(final_stack,raster_with_ndvi)
}


```


Load the boundary of the AOI and use it to mask the raster stack
This is needed to remove the black areas with no data... the dark region around the RGB plot above

```{r}
roi <-  st_read("/Users/rragankonywa/OneDrive/Moyodi_classification/Data/plan A/polygon project Area/PA.shp") %>% 
  st_transform(crs = crs(train_data))

final_stack <- mask(final_stack, roi)

```

Now extract raster values at each training point for each band in the final_stack

```{r}
train_df <- extract(final_stack, train_data, df = T, fun= mean)

```

Add the class and id variables from the original train csv(train_data) as columns in the extracted values (train_df)

```{r}
train_df_final <- train_df %>% 
  select(-ID) %>% 
  add_column(train_data %>% 
  select(id), .before="B_img1") %>% 
  select(-geometry)

train_df_final <- train_df_final[complete.cases(train_df_final),]

```

Fitting the model randomforest
```{r}

train_df_final$id <- factor(train_df_final$id)

# TODO: check proper settings of random forest algorithm
rfmodel <- randomForest(id ~., data = train_df_final, ntree =3000, nmtry = 400)
rfmodel
```

```{r}
predicted <- predict(final_stack, rfmodel)

plot(predicted)
```

```{r}
class_names <- train_data$Name %>% 
  unique()
class_id <- train_data$id %>% 
  unique()
#x <- as.data.frame(cbind(class_names,class_id))
classcolor <- c("#5475A8", "#B50000", "#D2CDC0", "#38814E", "#AF963C", "#D1D182", "#FBF65D", "#C8E6F8")
```


```{r}
predicted2 <- ratify(predicted)
predicted2 <- as.factor(predicted2)
rat <- levels(predicted2)[[1]]
rat[["zn"]] <- c(class_names)
levels(predicted2) <- rat

#rasterVis::levelplot(predicted2)

levelplot(predicted2, maxpixels = 1e6,
          col.regions = classcolor,
          scales=list(draw=FALSE),
          main = "Random Forest")

```

